---
title: "Part_1-4"
author: "Tianchuan Gao, Ruiqi Yan, Jiacheng Wu"
date: "5/4/2022"
output: 
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
library(ggplot2)
library(tidyverse)
library(MASS)
library(kableExtra)
theme_set(theme(legend.position = "bottom"))
```

# Hurrican Data
hurricane703.csv collected the track data of 703 hurricanes in  the North Atlantic area since 1950. For all the storms, their location (longitude \& latitude) and maximum wind speed were recorded every 6 hours. The data includes the following variables 

1. **ID**:  ID of the hurricanes
2. **Season**: In which \textbf{year} the hurricane occurred 
3. **Month**: In which \textbf{month} the hurricane occurred 
4. **Nature**:  Nature of the hurricane 
  + ET: Extra Tropical
  + DS: Disturbance
  + NR: Not Rated
  + SS: Sub Tropical
  + TS: Tropical Storm
5. **time**: dates and time of the record  
6. **Latitude** and **Longitude**:  The location of  a hurricane check point 
7. **Wind.kt**  Maximum wind speed (in Knot) at each check point 

## load and clean data

```{r}
load("df_rm.RData")
## create variables of interest
dt = df_rm %>% 
  janitor::clean_names() %>% 
  group_by(id) %>% 
  mutate(wind_lag = lag(wind_kt,1), ##previous wind speed
         lat_shift = lag(latitude, 1) - lag(latitude, 2), ## change of lat
         long_shift = lag(longitude, 1) - lag(longitude, 2), ## change of long
         wind_shift = lag(wind_kt, 1) - lag(wind_kt, 2)) %>%  ##change of wind speed
  drop_na()

## group by hurricane and create design matrix for each hurricane
df_mcmc = dt %>% 
  dplyr::select(id,wind_kt,wind_lag,lat_shift, long_shift, wind_shift) %>% 
  nest(y = wind_kt, x_matrix=wind_lag:wind_shift) %>% 
  mutate(x_matrix = map(.x = x_matrix, ~model.matrix(~., data = .x)),
         y = map(y, pull))

y_obs = df_mcmc$y ## list of wind speed for each hurricane
x_matrix = df_mcmc$x_matrix ## list of design matrix for each hurricane
n_obs = nrow(dt) ## number of

## list of observation data: wind speed and design matrix
obs_list = list(y_obs = y_obs,
           x_matrix = x_matrix)

n_freq = dt %>% group_by(id) %>% count() %>% pull(n) ## number of obs for each hurricane
d = 5  ## dimension of beta_i
m = 681  ## number of hurricanes
n_beta = m*d  ## total number of beta
n_theta = n_beta+d+1+15  ## total number of parameters
mu_ind = (n_beta+1):(n_beta+d) ## index of mu
sigma_ind = n_beta+d+1 ## index of sigma
cov_m_ind = (n_beta+d+2):n_theta ## index of elements of inverse covariance matrix
rep_time = rep(n_freq,d) ## repeat time for each beta
```


## Part I: Likelihood, Prior, Posterior and Conditional Posterior

```{r}
## log posterior for beta_j

log_posterior_beta = function(beta, sigma, mu, sigma_p,j){
  x = obs_list$x_matrix[[j]]
  y = obs_list$y_obs[[j]]
  return(-t(y - x %*% beta) %*% (y - x %*% beta)/(2*sigma)
         -(1/2)*t(beta-mu)%*%sigma_p%*%(beta-mu))
}

## log posterior for sigma

log_posterior_sigma = function(beta_frame, sigma){
  ## make sure sigma is positive
  if(sigma <= 0){
    return(-Inf)
  } else{
    log_lik = -(n_obs/2)*log(sigma)
    -sum((dt$wind_kt-beta_frame[,1]
          -dt$wind_lag*beta_frame[,2]
          -dt$lat_shift*beta_frame[,3]
          -dt$long_shift*beta_frame[,4]
          -dt$wind_shift*beta_frame[,5])^2/(2*sigma))
    log_prior = -log(sigma)
    return(log_lik + log_prior)
  }
}

## calculate the number of unique value
numunique = function(x){length(unique(x))}

trace_inverse = function(x){
  m = matrix(0,5,5)
  m[lower.tri(m, diag = TRUE)] = x[cov_m_ind]
  m[upper.tri(m)] = t(m)[upper.tri(m)]
  return(sum(diag(m)))
}
```

### Likelihood

$$
L(\boldsymbol Y \mid \boldsymbol \theta) \propto \prod_{i=1}^{m}\left(\sigma^{2}\right)^{-\frac{n_{i}}{2}} \exp \left\{-\frac{1}{2 \sigma^{2}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\}
$$
where $m$ is the number of hurricane and $n_i$ is the number of observations for $i^{th}$ hurricane


### Prior

$$
\pi\left(\boldsymbol\theta=(\boldsymbol B, \boldsymbol \mu, \Sigma^{-1}, \sigma^{2})\right) \propto\left(\sigma^{2}\right)^{-1} \left|\Sigma^{-1}\right|^{d+1} \exp \left\{-\frac{1}{2} \operatorname{tr}\left(\Sigma^{-1}\right)\right\}\prod_{i=1}^{m}\left|\Sigma^{-1}\right|^{\frac{1}{2}}\exp \left\{-\frac{1}{2}\left(\boldsymbol \beta_{i}-\boldsymbol \mu \right)^{T} \Sigma^{-1}(\boldsymbol \beta_{i}-\boldsymbol \mu)\right\}
$$
where $d$ is the dimension of $\boldsymbol \mu$


### Posterior

$$
\begin{aligned}
\pi(\boldsymbol \theta \mid \boldsymbol Y) & \propto\left(\sigma^{2}\right)^{-1}\left|\Sigma^{-1}\right|^{d+1} 
\exp \left\{-\frac{1}{2}+ \operatorname{tr}\left(\Sigma^{-1}\right)\right\}\\ 
&\times \prod_{i=1}^{m}\left(\sigma^{2}\right)^{-\frac{n_{i}}{2}}\left|\Sigma^{-1}\right|^{\frac{1}{2}} 
\exp \left\{-\frac{1}{2 \sigma^{2}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\} 
\exp \left\{-\frac{1}{2}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\right\}\\
&= \left(\sigma^{2}\right)^{-\left(1+\frac{\sum_{i=1}^{m} n_{i}}{2} \right)}\left|\Sigma^{-1}\right|^{d+1+\frac{m}{2}} 
\exp \left\{-\frac{1}{2} \operatorname{tr}\left(\Sigma^{-1}\right) \right\}
\exp \left\{-\frac{1}{2} \sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\right\} \\
&\times \exp \left\{-\frac{1}{2 \sigma^{2}} \sum_{i=1}^{m}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\}
\end{aligned}
$$


### Conditional Posterior

$$
\begin{aligned}
&\boldsymbol \beta_i:
\pi(\boldsymbol \beta_i \mid \boldsymbol \theta_{(-\boldsymbol \beta_i)} \boldsymbol Y) \propto \exp \left\{-\frac{1}{2}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)-\frac{1}{2 \sigma^{2}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\}\\
&\boldsymbol \mu:
\pi\left(\boldsymbol \mu \mid \boldsymbol \theta_{(-\boldsymbol \mu)}, \boldsymbol Y\right)\propto
\exp \left\{-\frac{1}{2} \sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\right\}
\end{aligned}
$$
$$
\begin{aligned}
(\boldsymbol \beta_{i}-\boldsymbol \mu)^{T} \Sigma^{-1}(\boldsymbol \beta_{i}-\boldsymbol \mu)
&=\operatorname{tr}\left(\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\right)\\
&=\operatorname{tr}\left(\Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)(\boldsymbol \beta_i-\boldsymbol \mu)^{T}\right)\\
\pi\left(\boldsymbol \mu \mid \boldsymbol \theta_{(-\boldsymbol \mu)}, Y\right)&\propto \exp\left\{-\frac{1}{2}\operatorname{tr}\left(\Sigma^{-1}\sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)(\boldsymbol \beta_i-\boldsymbol \mu)^{T}\right)\right\}\\
&= \exp\left\{-\frac{1}{2}\operatorname{tr}\left(\Sigma^{-1}m\left(\boldsymbol \mu - \bar{\boldsymbol \beta}\right)(\boldsymbol \mu - \bar{\boldsymbol \beta})^{T}\right)\right\}\\
&= \exp \left\{-\frac{1}{2}(\boldsymbol \mu-\bar{\boldsymbol \beta})^{T}\Sigma^{-1} m(\boldsymbol \mu-\bar{\boldsymbol \beta})\right\}\\
&\Rightarrow \boldsymbol \mu \mid \boldsymbol \theta_{(-\boldsymbol \mu)}, \boldsymbol Y \sim N(\bar{\boldsymbol \beta},\Sigma/m)
\end{aligned} 
$$
$$
\sigma^2: \pi\left(\sigma^{2} \mid \boldsymbol \theta_{(-\sigma^{2})}, \boldsymbol Y\right) \propto \left(\sigma^{2}\right)^{-\left(1+\frac{\sum_{i=1}^{m} n_{i}}{2} \right)}
\times \exp \left\{-\frac{1}{2 \sigma^{2}} \sum_{i=1}^{m}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\}
$$


$$
\begin{aligned}
\Sigma^{-1}:
\pi\left(\Sigma^{-1} \mid \boldsymbol \theta_{\left(-\Sigma^{-1}\right)},\boldsymbol Y\right)
&\propto
\left|\Sigma^{-1}\right|^{d+1+\frac{m}{2}} \exp \left\{-\frac{1}{2} \operatorname{tr}\left(\Sigma^{-1}\right)\right\}\exp \left\{-\frac{1}{2} \sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\right\}\\
&=\left|\Sigma^{-1}\right|^{d+1+\frac{m}{2}}\exp \left\{-\frac{1}{2}\operatorname{tr}\left(\Sigma^{-1}\right)-\frac{1}{2}\operatorname{tr}\left(\Sigma^{-1} \sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T}\right)\right\}\\
&=\left|\Sigma^{-1}\right|^{d+1+\frac{m}{2}}
\exp \left\{-\frac{1}{2}\operatorname{tr}\left(\Sigma^{-1}\left(I+\sum_{i=1}^{m}\left(\beta_{i}-\mu\right)\left(\beta_{1}-\mu\right)^{T}\right)\right)\right\}\\
&\Rightarrow \Sigma^{-1} \mid \boldsymbol \theta_{\left(-\Sigma^{-1}\right)}, \boldsymbol Y \sim \mathrm{Wishart}\left(3d+3+m,\left(\boldsymbol I+\sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T}\right)^{-1} \right)
\end{aligned}
$$


## Part II: MCMC Algorithm

We could generate full conditional posterior distribution for some parameters but not all, so we apply hybrid algorithm consisting with Metropolis-Hastings steps and Gibbs steps.


MH steps for $\beta_{ij}$:
Sampling proposed $\beta_{ij}^{\prime}$, $j = 0,1...4$ for $i^{th}$ hurricane from proposal distribution $U\left(\beta_{ij}^{(t)}-a_{ij},\beta_{ij}^{(t)}+a_{ij}\right)$, where $a_{ij}$ is the search window for $\beta_{ij}$. Since the proposals are symmetry, the accepting or rejecting the proposed $\beta_{ij}^{\prime}$ depends on the ratio of posterior distribution. Some of the parameters in $\boldsymbol \theta$ could be cancelled out, so the ratio simplified to be:

$$
\frac{\pi\left(\boldsymbol \beta_{i}^{\prime},\boldsymbol \theta_{(-\boldsymbol \beta_i)}^{(t)} \mid \boldsymbol Y\right)}{\pi\left(\boldsymbol \beta_{i}^{(t)}, \boldsymbol \theta_{(-\boldsymbol \beta_i)}^{(t)} \mid \boldsymbol Y\right)}
=\frac{\exp \left\{-\frac{1}{2 \sigma^{2(t)}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{\prime}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{\prime}\right)-\frac{1}{2} \left(\boldsymbol \beta_{i}^{\prime}-\boldsymbol \mu^{(t)}\right)^{T}\Sigma^{-1^{(t)}}\left(\boldsymbol \beta_{i}^{\prime}-\boldsymbol \mu^{(t)}\right)\right\}}{\exp \left\{-\frac{1}{2 \sigma^{2(t)}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t)}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t)}\right)-\frac{1}{2}\left(\boldsymbol \beta_{i}^{(t)}-\boldsymbol \mu^{(t)}\right){ }^{T} \Sigma^{-1(t)}\left(\boldsymbol \beta_{i}^{(t)}-\boldsymbol \mu^{(t)}\right)\right\}}
$$
where $\boldsymbol \beta_i^{\prime}$ consisting with $\beta_{ij}^{\prime}$, $\beta_{ik}^{(t)}$ for $k>j$ and $\beta_{ik}^{(t+1)}$ for $k<j$.

The log of the ratio:
$$
\begin{aligned}
\log \frac{\pi\left(\boldsymbol \beta_{i}^{\prime},\boldsymbol \theta_{(-\boldsymbol \beta_i)}^{(t)} \mid \boldsymbol Y\right)}{\pi\left(\boldsymbol \beta_{i}^{(t)}, \boldsymbol \theta_{(-\boldsymbol \beta_i)}^{(t)} \mid \boldsymbol Y\right)} &=-\frac{1}{2}\left(\frac{\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{\prime}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{\prime}\right)}{\sigma^{2(t)}}+\left(\boldsymbol \beta_{i}^{\prime}-\boldsymbol \mu^{(t)}\right)^{T} \Sigma^{-1^{(t)}}\left(\boldsymbol \beta_{i}^{\prime}-\boldsymbol \mu^{(t)}\right)\right)\\
&+\frac{1}{2}\left(\frac{\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t)}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t)}\right)}{\sigma^{2(t)}}+\left(\boldsymbol \beta_{i}^{(t)}-\boldsymbol \mu^{(t)}\right)^{T} \Sigma^{{-1}^{(t)}}\left(\boldsymbol \beta_{i}^{(t)}-\boldsymbol \mu^{(t)}\right)\right)
\end{aligned}
$$

Then we randomly sample $u$ from $U(0,1)$ and compare $\log(u)$ with the log ratio. If the $\log(u)$ is smaller, we accept $\beta_{ij}^{\prime} = \beta_{ij}^{(t+1)}$, otherwise we reject $\beta_{ij}^{\prime}$ and $\beta_{ij}^{(t)} = \beta_{ij}^{(t+1)}$.


Then, Gibb step for $\boldsymbol \mu$: Sample $\boldsymbol \mu^{(t+1)}$ from $N \left(\bar{\boldsymbol \beta}^{(t+1)} ,\Sigma^{(t)}/m\right)$, where $\bar{\boldsymbol \beta}^{(t+1)}$ is the average over $\boldsymbol \beta_i^{(t+1)}$.


Next, MH step to generate $\sigma^{2^{\prime}}$ from $U\left(\sigma^{2^{(t)}}-a_{\sigma^2},\sigma^{2^{(t)}}+a_{\sigma^2}\right)$. Firstly, check whether $\sigma^{2^{\prime}}$ is positive, if not, we reject $\sigma^{2^{\prime}}$. Then, we randomly sample $u$ from $U(0,1)$ and compare $\log(u)$ with
the log posterior ratio. The log posterior ratio
$$
\begin{aligned}
\log \frac{\pi\left(\sigma^{2^{\prime}},\boldsymbol \theta^{(t)}_{(-\sigma^{2})} \mid \boldsymbol Y\right)}{\pi\left( \sigma^{2^{(t)}},\boldsymbol \theta^{(t)}_{(-\sigma^{2})} \mid \boldsymbol Y\right)} 
&= -\left(1+\frac{M}{2}\right)\log(\sigma^{2^{\prime}})-\frac{1}{2\sigma^{2^{\prime}}}\sum_{i=1}^{m}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t+1)}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t+1)}\right)\\
&+\left(1+\frac{M}{2}\right)\log(\sigma^{2^{(t)}})+\frac{1}{2\sigma^{2^{(t)}}}\sum_{i=1}^{m}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t+1)}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t+1)}\right)
\end{aligned}
$$

where $M$ is total number of observation for all hurricanes. If the $\log(u)$ is smaller, we accept $\sigma^{2^{\prime}} = \sigma^{2^{(t+1)}}$.


Finally, we sample $\Sigma^{-1^{(t+1)}}$ from  $\mathrm{Wishart}\left(3d+3+m,\left(\boldsymbol I+\sum_{i=1}^{m}\left(\boldsymbol \beta_{i}^{(t+1)}-\boldsymbol \mu^{(t+1)}\right)\left(\boldsymbol \beta_{i}^{(t+1)}-\boldsymbol \mu^{(t+1)}\right)^{T}\right)^{-1} \right)$ 


```{r}
componentwisemixedMH=function(a, a_sigma, beta_0,sigma_0,cov_0, nrep=10000){
  sigma_p = cov_0
  beta = beta_0
  mu = colMeans(beta)
  sigma = sigma_0
  theta_chain_mix = matrix(0, nrow = nrep, ncol = n_theta)
  for(i in 1:nrep){
    sigma_m = solve(sigma_p)
    # update beta by block of each hurricane
    for(j in 1:m){
      for(k in 1:5){
        prop = beta[j,]
        prop[k] = prop[k]+2*(runif(1)-0.5)*a[(k-1)*681+j]
        if(log(runif(1)) < (log_posterior_beta(prop, sigma, mu, sigma_p,j) 
                            - log_posterior_beta(beta[j,], sigma, mu, sigma_p,j))){
          beta[j,k] = prop[k]
        }
      }
    }
    # update mu
    beta_mean = colMeans(beta)
    mu = mvrnorm(1,beta_mean,Sigma = sigma_m/m)
    
    # update sigma
    beta_frame = matrix(rep(c(beta),times = rep_time), ncol = d)
    prop = sigma
    prop = prop+2*(runif(1)-0.5)*a_sigma
    if(log(runif(1)) < (log_posterior_sigma(beta_frame, prop) 
                        - log_posterior_sigma(beta_frame, sigma))){
      sigma = prop
    }
    # update inverse covariance matrix
    beta_mu = t(beta)-mu
    S = solve(diag(d)+beta_mu%*%t(beta_mu))
    sigma_p = rWishart(1,df = 3*d+3+m,Sigma = S)
    sigma_p = apply(sigma_p,2,c)
    theta_chain_mix[i,] = c(c(beta),mu,sigma,sigma_p[lower.tri(sigma_p, diag = T)])
  }
  return(theta_chain_mix)
}
```


## Part III: Implementation and Estimations

### Iinitial Value

```{r}
## beta_i_0: MLR on ith hurricane and get the coefficient as beta_i_0
## mu_0: average over the beta_i_0 as the mu_0
## sigma_0: get the MSE of each MLR and take average as sigma_0
## sigma_m_0^-1: take the inverse of covariance matrix of beta_i_0 as the initial sigma_m_0^-1 

starting_beta = df_mcmc %>% 
  mutate(lm_fit = map2(.x = x_matrix, .y = y, ~lm(.y~.x[,-1], )),
         lm_0 = map_dbl(.x = lm_fit, ~coef(.x)[1]),
         lm_1 = map_dbl(.x = lm_fit, ~coef(.x)[2]),
         lm_2 = map_dbl(.x = lm_fit, ~coef(.x)[3]),
         lm_3 = map_dbl(.x = lm_fit, ~coef(.x)[4]),
         lm_4 = map_dbl(.x = lm_fit, ~coef(.x)[5]),
         sigma = map_dbl(.x = lm_fit, ~mean(.x$residuals^2))) %>% 
  ungroup() %>% 
  dplyr::select(lm_0:sigma)

sigma_0 = mean(starting_beta$sigma)

starting_beta = starting_beta%>%dplyr::select(-sigma) %>% as.matrix()
beta_0 = lm(wind_kt~wind_lag+lat_shift+long_shift+wind_shift, data = dt)
starting_beta[which(is.na(starting_beta[,2])),2] = coef(beta_0)[2]
starting_beta[which(is.na(starting_beta[,5])),5] = coef(beta_0)[5]

cov_0 = solve(cov(starting_beta))
```

Initial Values:

- $\boldsymbol \beta_{i}$ - Fit multivariate linear regression (MLR) for $i^{th}$ hurricane and use the coefficients as $\boldsymbol \beta_{i0}$

- $\boldsymbol \mu$ - Average over all $\boldsymbol \beta_{i0}$ as $\boldsymbol \mu_0$ 

- $\sigma^2$ - $\hat{\sigma}^2_i$ is the mean square residuals of the MLR model on $i^{th}$ hurricane. Take the mean over all $\hat{\sigma}^2_i$ as $\sigma^2_0$

- $\Sigma^{-1}$ - Generate the covaraiance matrix of $\boldsymbol \beta_{i0}$ and take the inverse of the matrix as $\Sigma^{-1}_0$

```{r}
## search window

ind_06 =  c(1,3,4,6,16,19,24,28,31,41,43,45,46,47,50,51,52,54,62,68,69,73,82,
            83,89,93:97,99,108,112,113,117:119, 121,122,125, 126,134,136,140,
            141,146,157, 158,163,182,206,217,225,239,249,250,255,261,263,274,
            282,297,306,310,315,332,333,335,344,345,355,357,371,372,381,397,
            399,400,403,404,412,415,416,418,421,427,431,433,436,439,442,445,
            448:450, 455,456,464,475,494,497,503,506,508,519,522:524,531,532,
            538,540,541,544,550,563,564,569,571,584,589,591,601,613,614,618,
            639,642,646,663,664)

ind_12 = c(8,13,38,57,63,64,66,72,80,84,114,115,137,145,149,159,164,165,186,
           196,204,207,218,223,234,236,238,243,246,251,254,260,266,268,275,
           283,288,290,291,294,307,323:326,330,334,336,354,368,370,377,378,
           380,387,395,398,401,428,429,432,437,454,457,458,460,469,480,482,
           487,488,491,495,499,504,510,517,526,527,534,537,539,545,547,551,
           570,572,575,583,595,596,599,602,603,609,610,612,619,624,630,634,
           637,640,641,648,660,665,666,671)

a2 = rep(0.08,m)
a2[ind_06] = 0.04
a2[ind_12] = 0.1
a3 = rep(1, m)
a3[203] = 0.8
a4 = rep(0.6,m)
a4[c(4,18,20,32,40,50,62,68,94,106,118,124,134,140,166,216,220,250,272,274,282,
     306,328,364, 376, 400, 412, 420, 434, 440, 442, 452, 456, 464, 508, 514, 
     522, 524, 562, 564, 568,584, 614, 642,646)] = 0.5

a5 = rep(0.3,m)
ind_14 = c(3,6,12,15,18,21,24,30,36,39,42 ,48, 57 ,63,66,72,75,78,81,84,87,93,
           102, 108,114,120,126,132,135,138,147,153,159,162,165,168,174,177,180,
           183,186,189,192,195,198,201,204,207,210,216,219,222,225,228,231,234,
           237,240,243,246,252,258,261,264,267,270,276,279,288,291,294,297,300,
           303,309,312,315,318,324,327,330,333,336, 339,342, 348, 351,354,357,
           360,363,366, 369, 372, 375, 378, 381, 387, 390, 393, 399, 402,405, 
           408, 411, 417, 420, 423, 426, 429,432, 435,438, 441, 444, 447, 450,
           453, 459, 462, 465, 468, 471, 474, 477, 480, 486, 489, 492,495,498, 
           501,504, 507, 510, 513, 525, 528, 534, 537, 540, 543, 546, 549, 555, 
           558, 564, 567, 570, 573,576, 579, 582, 585,588, 594, 600, 603, 606, 
           609, 612, 615, 624, 630, 633, 636, 639, 642,645, 648, 651, 654, 657, 
           660, 663, 666,669, 672, 675, 678, 681)

a5[ind_14] = 0.4

a = c(rep(1.1,m),a2,a3,a4,a5)
```

### Implement 10000 MCMC iterations

```{r eval=FALSE, echo = TRUE}
set.seed(1971)
theta_chain = componentwisemixedMH(a, a_sigma = 2, beta_0 = starting_beta, 
                                   sigma_0, cov_0, nrep=10000)
```


### Search Window and Acceptance Rate

```{r}
load("mcmc_chain_1.RData")
load("mcmc_chain_2.RData")
load("mcmc_chain_3.RData")
load("mcmc_chain_4.RData")
load("mcmc_chain_5.RData")
load("mcmc_chain_6.RData")

theta_chain = rbind(theta_chain_1, theta_chain_2, theta_chain_3, 
                    theta_chain_4, theta_chain_5, theta_chain_6)

a_rate = apply(theta_chain, MARGIN = 2, numunique)/10000

a_rate_matrix = matrix(a_rate[1:n_beta], ncol = 5)

a_range = range(c(a_rate[1:n_beta], a_rate[sigma_ind]))

a_0_range = range(a_rate_matrix[,1])*100 # search window 1.1
a_1_range = range(a_rate_matrix[,2])*100 # search window: 0.04-0.1
a_2_range = range(a_rate_matrix[,3])*100 # search window: 0.8-1
a_3_range = range(a_rate_matrix[,4])*100 # search window:0.5-0.6
a_4_range = range(a_rate_matrix[,5])*100 # search window: 0.4-0.5
a_sigma = a_rate[sigma_ind]*100
```

There is MH step in the MCMC algorithm so the choice of search window is important. The acceptance rate of the MH step is around `r a_range[1]` to `r a_range[2]`, so the search window of our algorithm is appropriate. We tune the search windows multiple times to achieve this result. Table 1 demonstrates the range of search windows $a$ and associated acceptance rates for $\boldsymbol \beta$ and $\sigma^2$.

```{r}
df_a = tibble(parameter = c("$\\boldsymbol{\\beta}_0$", "$\\boldsymbol{\\beta}_1$", 
                            "$\\boldsymbol{\\beta}_2$", "$\\boldsymbol{\\beta}_3$", 
                            "$\\boldsymbol{\\beta}_4$", "$\\boldsymbol{\\sigma}^2$"),
              searchwindow = c("1.1",
                                "0.04-0.1",
                                "0.8-1",
                                "0.5-0.6",
                                "0.4-0.5",
                                "2"),
              acceptancerate = c(paste0(a_0_range[1],"-",a_0_range[2]),
                                  paste0(a_1_range[1],"-",a_1_range[2]),
                                  paste0(a_2_range[1],"-",a_2_range[2]),
                                  paste0(a_3_range[1],"-",a_3_range[2]),
                                  paste0(a_4_range[1],"-",a_4_range[2]),
                                  paste0(a_sigma)))

label_2 = c("","search window", "acceptance rate ($\\%$)")

df_a %>% kable(format = "latex",escape=FALSE, booktabs = TRUE, col.names = label_2,
               centering = T, vline = "|", 
               linesep = c("\\addlinespace", "\\addlinespace",
                           "\\addlinespace", "\\addlinespace",
                           "\\addlinespace","\\addlinespace"),
               caption = "Range of Search Window and Acceptance Rate for paraemters used MH step") %>% 
  kable_styling(latex_options = "hold_position")
```
 
```{r}
theta_trace_inverse = apply(theta_chain, 1, trace_inverse)
```
### Convergence trace plots

```{r}
L = 0.7
cex = 0.6
cex_lab = 0.9
par(mfrow=c(3,4),mgp = c(0.5,0.1,0), tck = -0.01)
par(mar = c(2, 2, 2, 0.3))
plot(theta_chain[,2], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(beta[0*2]), line = L, cex.lab = cex_lab)
plot(theta_chain[,m+2], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(beta[12]), line = L, cex.lab = cex_lab)
mtext("                                 Covergence Plots of Selected Parameters", side = 3, line = 0.6)
plot(theta_chain[,m*2+2], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(beta[22]), line = L, cex.lab = cex_lab)
plot(theta_chain[,m*3+2], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(beta[32]), line = L, cex.lab = cex_lab)
plot(theta_chain[,m*4+2], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(beta[42]), line = L, cex.lab = cex_lab)
plot(theta_chain[,3406], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(mu[0]), line = L, cex.lab = cex_lab)
plot(theta_chain[,3407], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(mu[1]), line = L, cex.lab = cex_lab)
plot(theta_chain[,3408], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(mu[2]), line = L, cex.lab = cex_lab)
plot(theta_chain[,3409], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(mu[3]), line = L, cex.lab = cex_lab)
plot(theta_chain[,3410], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(mu[4]), line = L, cex.lab = cex_lab)
plot(theta_chain[,3411], type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression(sigma^2), line = L, cex.lab = cex_lab)
plot(theta_trace_inverse, type = "l", cex.axis = cex, xlab = "", ylab = "")
abline(v = 5500, lty = 2, col = 2)
title(ylab = expression("Trace of "*Sigma^-1), line = L, cex.lab = cex_lab)
```

### Histograms

```{r}
L = 1
cex = 0.7
par(mfrow=c(3,4))
par(mar = c(3, 1, 3, 1), mgp = c(0.1,0.1,-0.2),tck = -0.01)
hist(theta_chain[5501:10000,2], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
title(cex.lab = cex_lab, xlab = expression(beta[0*2]), line = L)
hist(theta_chain[5501:10000,m+2], nclass = 70, main = NULL, xlab = "", ylab = "", cex.axis = cex)
mtext("                                 Histograms of Selected Parameters", side = 3, line = 0.6)
title(cex.lab = cex_lab, xlab = expression(beta[1*2]), line = L)
hist(theta_chain[5501:10000,m*2+2], nclass = 50, main = NULL, cex.axis = cex, xlab = "", ylab = "")
title(cex.lab = cex_lab, xlab = expression(beta[2*2]), line = L)
hist(theta_chain[5501:10000,m*3+2], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
title(cex.lab = cex_lab, xlab = expression(beta[3*2]), line = L)
hist(theta_chain[5501:10000,m*4+2], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
title(cex.lab = cex_lab, xlab = expression(beta[4*2]), line = L)
hist(theta_chain[5501:10000,3406], nclass = 60, main = NULL, xlab = "", ylab = "", cex.axis = cex)
title(cex.lab = cex_lab, xlab = expression(mu[0]), line = L)
hist(theta_chain[5501:10000,3407], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
title(cex.lab = cex_lab, xlab = expression(mu[1]), line = L)
hist(theta_chain[5501:10000,3408], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
title(cex.lab = cex_lab, xlab = expression(mu[2]), line = L)
hist(theta_chain[5501:10000,3409], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
title(cex.lab = cex_lab, xlab = expression(mu[3]), line = L)
hist(theta_chain[5501:10000,3410], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
title(cex.lab = cex_lab, xlab = expression(mu[4]), line = L)
hist(theta_chain[5501:10000,3411], nclass = 50, main = NULL, xlab = "", ylab  = "", cex.axis = cex)
title(cex.lab = cex_lab, xlab = expression(sigma^2), line = L)
hist(theta_trace_inverse, nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
title(cex.lab = cex_lab, xlab = expression("Trace of "*Sigma^-1), line = L)
```

### Estimates (burn-in first 5500 runs)

```{r}
## take average over 5501-10000 as estimates
bs_estimates = unname(apply(theta_chain[5501:10000,], 2, mean))


inv_cov_m = matrix(0,5,5)
inv_cov_m[lower.tri(inv_cov_m, diag = TRUE)] = bs_estimates[cov_m_ind]
inv_cov_m[upper.tri(inv_cov_m)] = t(inv_cov_m)[upper.tri(inv_cov_m)]


cov_m = solve(inv_cov_m)

## estimates
sigma = bs_estimates[3411]
mu = bs_estimates[3406:3410]
beta = matrix(bs_estimates[1:n_beta], ncol = d)
```

```{r}
cex = 0.8
cex_lab = 1
par(mfrow=c(2,3),mgp = c(0.1,0.1,-0.3),tck = -0.01)
par(mar = c(3, 2, 3, 2))
hist(beta[,1], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
abline(v = mu[1], lty = 2, col = 2)
title(xlab = expression(beta[0]), cex.lab = cex_lab, line = L)
hist(beta[,2], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
abline(v = mu[2], lty = 2, col = 2)
title(xlab = expression(beta[1]), cex.lab = cex_lab, line = L)
mtext(expression("Histograms of estimated "*hat(beta)*" of All Hurricanes"), side = 3, line = 0.6)
hist(beta[,3], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
abline(v = mu[3], lty = 2, col = 2)
title(xlab = expression(beta[2]), cex.lab = cex_lab, line = L)
hist(beta[,4], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
abline(v = mu[4], lty = 2, col = 2)
title(xlab = expression(beta[3]), cex.lab = cex_lab, line = L)
hist(beta[,5], nclass = 50, main = NULL, xlab = "", ylab = "", cex.axis = cex)
abline(v = mu[5], lty = 2, col = 2)
title(xlab = expression(beta[4]), cex.lab = cex_lab, line = L)
```

### Chart of the estimates of $\mu$, $\sigma^2$ and covaraince matrix

```{r}
sigmap_m = round(cov_m[lower.tri(cov_m, diag = T)],3)
output1 = t(rbind(round(mu[1], 2), round(mu[2], 2), round(mu[3], 2), round(mu[4], 2), round(mu[5], 2), round(sigma, 2),sigmap_m[1],sigmap_m[6], sigmap_m[10], sigmap_m[13], sigmap_m[15]))
colnames(output1) = c("$\\mu_0$", "$\\mu_1$", "$\\mu_2$", "$\\mu_3$", "$\\mu_4$", "$\\sigma^2$","$\\sigma_{00}^2$","$\\sigma_{11}^2$","$\\sigma_{22}^2$","$\\sigma_{33}^2$","$\\sigma_{44}^2$")
rownames(output1) = "Estimates"
output1 %>% knitr::kable(caption = "Bayesian Estimates for $\\boldsymbol \\mu$ and $\\sigma^2$", escape = FALSE) %>% 
  kable_styling(latex_options = "hold_position")
```

```{r}
sigmap = round(bs_estimates[3412:3426],2)
cor = cov2cor(cov_m)
sigmap_c = round(cor[lower.tri(cor, diag = T)],3)
```

$$
\hat{\Sigma}^{-1} = 
\begin{bmatrix}
`r sigmap[1]` & `r sigmap[2]` & `r sigmap[3]` & `r sigmap[4]` & `r sigmap[5]`\\
`r sigmap[2]` & `r sigmap[6]` & `r sigmap[7]` & `r sigmap[8]` & `r sigmap[9]`\\
`r sigmap[3]` & `r sigmap[7]` & `r sigmap[10]` & `r sigmap[11]` & `r sigmap[12]`\\
`r sigmap[4]` & `r sigmap[8]` & `r sigmap[11]` & `r sigmap[13]` & `r sigmap[14]`\\
`r sigmap[5]` & `r sigmap[9]` & `r sigmap[12]` & `r sigmap[14]` & `r sigmap[15]`
\end{bmatrix} \text{ }
\hat{\rho} = 
\begin{bmatrix}
`r sigmap_c[1]` & `r sigmap_c[2]` & `r sigmap_c[3]` & `r sigmap_c[4]` & `r sigmap_c[5]`\\
`r sigmap_c[2]` & `r sigmap_c[6]` & `r sigmap_c[7]` & `r sigmap_c[8]` & `r sigmap_c[9]`\\
`r sigmap_c[3]` & `r sigmap_c[7]` & `r sigmap_c[10]` & `r sigmap_c[11]` & `r sigmap_c[12]`\\
`r sigmap_c[4]` & `r sigmap_c[8]` & `r sigmap_c[11]` & `r sigmap_c[13]` & `r sigmap_c[14]`\\
`r sigmap_c[5]` & `r sigmap_c[9]` & `r sigmap_c[12]` & `r sigmap_c[14]` & `r sigmap_c[15]`
\end{bmatrix}
$$

```{r}
#
#\hat{\Sigma} = 
#\begin{bmatrix}
#`r sigmap_m[1]` & `r sigmap_m[2]` & `r sigmap_m[3]` & `r sigmap_m[4]` & `r sigmap_m[5]`\\
#`r sigmap_m[2]` & `r sigmap_m[6]` & `r sigmap_m[7]` & `r sigmap_m[8]` & `r sigmap_m[9]`\\
#`r sigmap_m[3]` & `r sigmap_m[7]` & `r sigmap_m[10]` & `r sigmap_m[11]` & `r sigmap_m[12]`\\
#`r sigmap_m[4]` & `r sigmap_m[8]` & `r sigmap_m[11]` & `r sigmap_m[13]` & `r sigmap_m[14]`\\
#`r sigmap_m[5]` & `r sigmap_m[9]` & `r sigmap_m[12]` & `r sigmap_m[14]` & `r sigmap_m[15]`
#\end{bmatrix}
```


## Part IV: Model Performance

```{r}
dt_res = dt %>% 
  dplyr::select(id, nature, season, month, wind_kt,wind_lag,lat_shift, long_shift, wind_shift) %>%
  nest(data = nature:wind_shift) %>% 
  ungroup() %>% 
  mutate(beta_0 = beta[,1],
         beta_1 = beta[,2],
         beta_2 = beta[,3],
         beta_3 = beta[,4],
         beta_4 = beta[,5]) %>% 
  unnest(data) %>% 
  mutate(wind_kt_pred = beta_0+beta_1*wind_lag+beta_2*lat_shift+beta_3*long_shift+beta_4*wind_shift) %>%
  group_by(id) %>% 
  mutate(r_square = (1-sum((wind_kt_pred-wind_kt)^2)/(sum((wind_kt-mean(wind_kt))^2))),
         n_obs = n(),
         adj_r_square = 1-(1-r_square)*(n_obs-1)/(n_obs-5),
         p_value = 1-pchisq(sum((wind_kt_pred-wind_kt)^2)/sigma, n_obs))
  
```


```{r}
r_square = (1-sum((dt_res$wind_kt_pred-dt_res$wind_kt)^2)/sum((dt_res$wind_kt-mean(dt_res$wind_kt))^2))
adj_r_square = 1-(1-r_square)*(n_obs-1)/(n_obs-n_beta)
```

The overall adjusted $R^2$ of the estimated Bayesian model is `r adj_r_square`.

```{r}
dt_rsquare = dt_res %>% 
  dplyr::select(id, r_square, adj_r_square, p_value, n_obs) %>% 
  distinct() %>% 
  arrange(n_obs)

n_lowp = dt_rsquare %>% filter(p_value < 0.05) %>% 
  nrow()

d_high = dt_rsquare %>% 
  filter(adj_r_square >=0.6 & adj_r_square <= 1) %>% 
  nrow()

d_moderate = dt_rsquare %>% 
  filter(adj_r_square >= 0.2 & adj_r_square< 0.6) %>% 
  nrow()

d_low = dt_rsquare %>% 
  filter(adj_r_square < 0.2) %>% 
  nrow()


label_1 = c("$R_{adj}^2$", "Count of Hurricanes","Percentage($\\%$)")

tibble(adj_r = c("0.6-1", "0.2-0.6","< 0.2"),
       count = c(d_high, d_moderate, d_low),
       prop = c(round(d_high/681,3)*100, round(d_moderate/681,3)*100, 
                round(d_low/681,3)*100)) %>% 
  kable(escape = F, col.names = label_1, 
        caption = "$R_{adj}^2$ for each hurricane") %>% 
  kable_styling(latex_options = "hold_position")

```

Furthermore, we check the model performance of individual hurricane. Even though most of the estimated models track hurricanes well with great adjusted $R^2$, some estimated models track the hurricanes extremely bad. `r round((d_low+d_moderate)/681,3)*100`% of the estimated models do not track the hurricane well and have adjusted $R^2$ less than 0.6. Few models have negative adjusted $R^2$. Table 2 show the distribution of the adjusted $R^2$ of the estimated Baysian models for individual hurricane.


```{r}
## visualize some hurricane prediction
dt_rsquare %>% 
  ggplot(aes(x = n_obs, y = adj_r_square))+
  geom_point()+
  geom_vline(xintercept = 40, linetype = 2)+
  geom_hline(yintercept = 0.5, linetype = 2)+
  geom_vline(xintercept = 20, linetype = 1)+
  geom_hline(yintercept = 0, linetype = 1)


dt_res %>% filter(id %in% c("DOG.1950","JANET.1955","FRANCES.1986",
                            "BARBARA.1954","GINGER.1967", "ARLENE.1971",
                            "NICOLE.2010","JERRY.2007","TEN.2005")) %>% 
  ungroup() %>% 
  mutate(adj_r_square = round(adj_r_square,3),
         id = factor(paste0(id, "(",adj_r_square,")")),
         id = reorder(id, adj_r_square)) %>% 
  group_by(id) %>% 
  mutate(time = 1:n()) %>%
  ggplot(aes(y = wind_kt, x = time))+
  geom_line(aes(color = "observation"))+
  geom_line(aes(y = wind_kt_pred, color = "prediction"))+
  scale_colour_manual(name = NULL,
values = c( "observation" = 1, "prediction" = 2))+
  facet_wrap(.~id, nrow = 3, scales = "free")+
  labs(x = "time",
       y = "wind speed",
       title = "Observed Wind Speed vs. Baysian Model Estimates")
```


We also perform the goodness-of-fit test for individual estimated model. We used residuals of Bayesian estimates, $r_{ij}$, of the $j^{th}$ observation in $i^{th}$ hurricane, to calculate the test statistics. $\chi^2_{stat} = \frac{\sum_{j=1}^{n_i}r_{ij}^2}{\sigma^2}$, where $\sigma^2$ is the estimate $\sigma^2$ from MCMC. Based on the normal assumption in intro, $\chi^2_{stat}\sim\chi^2_{n_i}$, where $n_i$ is the number of observation for $i^{th}$ hurricane. The estimated models of `r n_lowp` hurricanes have p-value less than 0.05, implying that those models do not fit well. 





