---
title: |
  | P8160 - Project 3
  | Baysian modeling of hurrican trajectories
author: |
  | Group 6
  |
  | Amy Pitts, Jiacheng Wu, Jimmy Kelliher, Ruiqi Yan & Tianchuan Gao
date: "2022-05-09"
output:
  pdf_document: default
  
header-includes:
  - \usepackage{amsmath,amssymb,amsthm,amsfonts,bbm,graphics,enumitem}
  - \usepackage{geometry,mathrsfs}
  - \usepackage{tikz}
  - \usepackage{pgfplots}
  - \usetikzlibrary{shapes,arrows}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyverse)
library(data.table)
library(maps)
library(gtsummary)

dt = read.csv("hurrican703.csv")
```


\vspace{75mm}

\begin{center}
{\bf Abstract}
\end{center}

|        abstract 

\newpage
# 1. Introduction

## 1.1. Background


## 1.2. Objectives





# 2. Methods

## 2.1. Data Cleaning and Exploratory Analysis

```{r message=FALSE, warning=TRUE, include=FALSE}
temp = table(dt$ID) %>% as.data.frame() 
```

There are `r length(unique(dt$ID))` unique hurricanes in this dataset all the occurred in the north american region between the years `r range(dt$Season)[1]` to `r range(dt$Season)[2]`. Data on the storm's location (longitude & latitude) and maximum wind speed were recorded every 6 hours. The number of observations we have for each storm range from `r range(temp$Freq)[1]` to `r range(temp$Freq)[2]` with a mean value of `r round(mean(temp$Freq))` observations per hurricane. Data is also collected on the storms month, and the nature of the hurricane; (Extra Tropical (ET), Disturbance (DS), Not Rated (NR), Sub Tropical (SS), and Tropical Storm (TS)). 



# 3. Results 



# 4. Discussion 

## 4.1. Summary of Findings


## 4.2. Limitations



## 4.3 Future Work



## 4.4. Group Contributions



\newpage

# References 

\begin{enumerate}[label={[\arabic*]}]
  \item  one 
  \item  two 
\end{enumerate}

\newpage 

# Appendices








