---
title: |
  | P8160 - Project 3
  | Baysian modeling of hurrican trajectories
author: |
  | Group 6
  |
  | Amy Pitts, Jiacheng Wu, Jimmy Kelliher, Ruiqi Yan & Tianchuan Gao
date: "2022-05-09"
output:
  pdf_document: default
  
header-includes:
  - \usepackage{amsmath,amssymb,amsthm,amsfonts,bbm,graphics,enumitem}
  - \usepackage{geometry,mathrsfs}
  - \usepackage{tikz}
  - \usepackage{pgfplots}
  - \usetikzlibrary{shapes,arrows}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


source("report_r_code.R")

theme_gtsummary_compact()
```


\vspace{75mm}

\begin{center}
{\bf Abstract}
\end{center}

|        abstract 

\newpage
# 1. Introduction

## 1.1. Background

Hurricanes are tropical storms that reach a wind speed of 74 miles (119 kilometers) per hour or greater. They also bring heavy rain, thunder storms, and other severe meteorological phenomena. 
Throughout history, hurricanes have brought about huge death tolls and financial loss, especially in coastal cities and countries. Therefore, it is crucial for researchers to predict the behaviors of hurricanes, take appropriate counter measures, and minimize casualty. 

## 1.2. Objectives

For this project, we aim to examine the factors that could predict the wind speed of hurricanes, using the track data of 703 hurricanes in the North Atlantic area since 1950. The data includes each hurricane’s geographical coordinates, maximum speeds recorded every 6 hours, the year (season) and month in which the hurricane took place, and its nature. We build a Bayesian model for each hurricane that predicts its wind speed at a given time based on its last recorded speed, and the change in longitude, latitude, and speed from the last observation. After deriving the posterior distributions, we use the Markov Chain Monte-Carlo Method to estimate the parameters. We then explore whether the season, month, and nature of the hurricane affect the wind speed and the impact of wind speed on death counts and financial loss. 



# 2. Methods

## 2.1. Data Cleaning and Exploratory Analysis

```{r message=FALSE, warning=TRUE, include=FALSE}
temp = table(dt$ID) %>% as.data.frame() 
```

There are `r length(unique(dt$ID))` unique hurricanes in this dataset all the occurred in the north american region between the years `r range(dt$Season)[1]` to `r range(dt$Season)[2]`. Data on the storm's location (longitude & latitude) and maximum wind speed were recorded every 6 hours. The number of observations we have for each storm range from `r range(temp$Freq)[1]` to `r range(temp$Freq)[2]` with a mean value of `r round(mean(temp$Freq))` observations per hurricane. Data is also collected on the storms month, and the nature of the hurricane; (Extra Tropical (ET), Disturbance (DS), Not Rated (NR), Sub Tropical (SS), and Tropical Storm (TS)). 

To conduct our analysis we require at least 7 observations for each unique hurricane. In addition, we are only concerned about observations that occurred on 6 hour intervals. The dataset includes a couple observations between the 6 hour time periods. For our analysis we are only going to include observations that are recorded on hour 0, 6, 12, and 18. In addition we will exclude all hurricane IDs that have less then 7 observations. Through this process we remove `r (dim(dt)[1]-dim(df_rm)[1])` observations so we are left with `r dim(df_rm)[1]` observations and `r length(unique(df_rm$ID))` unique hurricanes. In addition we also created variables of lag difference ($t-6$ to $t-12$) for latitude, longitude and wind speed as $\Delta_{i,1}(t),\Delta_{i,2}(t), \Delta_{i,3}(t)$ and lag of wind speed as $Y_{i}(t-6)$. 

\begin{center}
\textit{Table 1: Data Characteristics shown by the Nature of the Hurricane variable}
\end{center}

```{r echo=FALSE, message=FALSE, warning=FALSE}
df_rm %>%
  #arrange(match(Month, month.name)) %>% 
  select(Nature, Wind.kt, Latitude, Longitude,  Month ) %>% 
  tbl_summary(by =Nature ) %>%
  add_overall() %>% 
  modify_spanning_header(c("stat_1", "stat_2","stat_3","stat_4","stat_5") ~ "**Nature of the Hurricane**") 

```


## 2.2 Bayesian Model for Hurricane Trajectories

Climate researchers are interested in modeling the hurricane trajectories to forecast the winds peed. To model the wind speed of the $i^{th}$ hurricane at time $t$ we will use
$$
Y_i(t) = \beta_{0,i}+\beta_{1,i}Y_i(t-6)+\beta_{2,i}\Delta_{i,1}(t)+\beta_{3,i}\Delta_{i,2}(t)+\beta_{4,i}\Delta_{i,3}(t) + \epsilon_i(t)
$$
Where  $\Delta_{i,1}(t)$, $\Delta_{i,2}(t)$ and $\Delta_{i,3}(t)$ are changes in latitude longitude and wind speed respectively between $t-6$ and $t$. The $\epsilon_i(t) \sim \mathcal{N}(0, \sigma^2)$ are independent across $t$. Let $\beta_i = ( \beta_{0,i},  \beta_{1,i},  \beta_{2,i}, \beta_{3,i}, \beta_{4,i})^T \sim \mathcal{N}(\mu, \Sigma)$ be multivariate normal distribution where $\mu \in \mathbb{R}^d$ and $\Sigma \in \mathbb{R}^{d \times d}$. For this model we will be assuming non-informative or weak prior distributions for our unknown parameters $\sigma^2$,  $\mu$ and $\Sigma$ . 
\[
\pi(\sigma^2) \propto \frac{1}{\sigma^2}, \ \ \ \ \ \ \pi(\mu) \propto 1, \ \ \ \ \ \  \pi\left(\Sigma^{-1}\right) \propto \left|\Sigma \right|^{-(d+1)}\exp\left\{ -\frac{1}{2} \Sigma^{-1}\right\} 
\]
Our goal is to estimate $\Theta = (\boldsymbol B, \boldsymbol \mu, \Sigma^{-1}, \sigma^{2})$. To do this we need to establish our likelihood and prior functions. Since this is a Bayesian model we have that the likelihood will be expressed as 
\[ 
L(\boldsymbol Y \mid \boldsymbol \Theta) \propto \prod_{i=1}^{m}\left(\sigma^{2}\right)^{-\frac{n_{i}}{2}} \exp \left\{-\frac{1}{2 \sigma^{2}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\} 
\]
where $m$ is the number of hurricane and $n_i$ is the number of observations for $i^{th}$ hurricane. The joint prior distribution is expressed as 
\[
\pi\left(\boldsymbol\Theta\right) \propto\left(\sigma^{2}\right)^{-1} \left|\Sigma^{-1}\right|^{d+1} \exp \left\{-\frac{1}{2} \operatorname{tr}\left(\Sigma^{-1}\right)\right\}\prod_{i=1}^{m}\left|\Sigma^{-1}\right|^{\frac{1}{2}}\exp \left\{-\frac{1}{2}\left(\boldsymbol \beta_{i}-\boldsymbol \mu \right)^{T} \Sigma^{-1}(\boldsymbol \beta_{i}-\boldsymbol \mu)\right\}
\]
where $d$ is the dimension of $\boldsymbol \mu$. Using the likelihood and priors derived above we can calculate the posterior distribution. 
$$
\begin{aligned}
\pi(\boldsymbol \Theta \mid \boldsymbol Y) & \propto\left(\sigma^{2}\right)^{-1}\left|\Sigma^{-1}\right|^{d+1} 
\exp \left\{-\frac{1}{2}+ \operatorname{tr}\left(\Sigma^{-1}\right)\right\}\\ 
&\times \prod_{i=1}^{m}\left(\sigma^{2}\right)^{-\frac{n_{i}}{2}}\left|\Sigma^{-1}\right|^{\frac{1}{2}} 
\exp \left\{-\frac{1}{2 \sigma^{2}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\} 
\exp \left\{-\frac{1}{2}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\right\}\\
&= \left(\sigma^{2}\right)^{-\left(1+\frac{\sum_{i=1}^{m} n_{i}}{2} \right)}\left|\Sigma^{-1}\right|^{d+1+\frac{m}{2}} 
\exp \left\{-\frac{1}{2} \operatorname{tr}\left(\Sigma^{-1}\right) \right\}
\exp \left\{-\frac{1}{2} \sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\right\} \\
&\times \exp \left\{-\frac{1}{2 \sigma^{2}} \sum_{i=1}^{m}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\}
\end{aligned}
$$

Due to the number of unknown parameters and the complexity we will need to use method to approximate this distribution. Before describing that process we first need to calculate the conditional posterior distribution for each unknown parameter. 
$$
\begin{aligned}
&\boldsymbol \beta_i:
\pi(\boldsymbol \beta_i \mid \boldsymbol \Theta_{(-\boldsymbol \beta_i)} \boldsymbol Y) \propto \exp \left\{-\frac{1}{2}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)-\frac{1}{2 \sigma^{2}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\}\\
&\boldsymbol \mu:
\pi\left(\boldsymbol \mu \mid \boldsymbol \Theta_{(-\boldsymbol \mu)}, \boldsymbol Y\right)  \sim N(\bar{\boldsymbol \beta},\Sigma/m) \\
&\sigma^2: \pi\left(\sigma^{2} \mid \boldsymbol \Theta_{(-\sigma^{2})}, \boldsymbol Y\right) \propto \left(\sigma^{2}\right)^{-\left(1+\frac{\sum_{i=1}^{m} n_{i}}{2} \right)}
\times \exp \left\{-\frac{1}{2 \sigma^{2}} \sum_{i=1}^{m}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\} \\
&\Sigma^{-1}:
\pi\left(\Sigma^{-1} \mid \boldsymbol \Theta_{\left(-\Sigma^{-1}\right)},\boldsymbol Y\right) \sim \mathrm{Wishart}\left(3d+3+m,\left(\boldsymbol I+\sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T}\right)^{-1} \right)
\end{aligned}
$$

To see a more detailed description of each conditional posterior please see Appendix A. 








## 2.3 Markov Chain Monte Carlo (MCMC) Algorithm
Due to the complexity of the above posterior distribution we will use a Markov chain Monte Carlo (MCMC) process. Since we could generate full conditional posterior distribution for some parameters but not all we will instead apply hybrid algorithm consisting with Metropolis-Hastings (MH) steps and Gibbs steps. We will preform a MH step for $\beta_{j,i}$, $\sigma^2$ and sample from the $\Sigma^{-1^{(t+1)}}$ distribution. We will also describe a gibbs step for $\mu$ using the  $\beta_{j,i}$ gathered. 


First is the MH steps for $\beta_{j,i}$. Sampling proposed $\beta_{j,i}^{\prime}$, $j = 0,1...4$ for $i^{th}$ hurricane from proposal distribution $U\left(\beta_{j,i}^{(t)}-a_{j,i},\beta_{j,i}^{(t)}+a_{j,i}\right)$, where $a_{j,i}$ is the search window for $\beta_{j,i}$. Since the proposals are symmetry, the accepting or rejecting the proposed $\beta_{j,i}^{\prime}$ depends on the ratio of posterior distribution. Some of the parameters in $\boldsymbol \Theta$ could be cancelled out, so the ratio simplified to be:

$$
\frac{\pi\left(\boldsymbol \beta_{i}^{\prime},\boldsymbol \Theta_{(-\boldsymbol \beta_i)}^{(t)} \mid \boldsymbol Y\right)}{\pi\left(\boldsymbol \beta_{i}^{(t)}, \boldsymbol \Theta_{(-\boldsymbol \beta_i)}^{(t)} \mid \boldsymbol Y\right)}
=\frac{\exp \left\{-\frac{1}{2 \sigma^{2(t)}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{\prime}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{\prime}\right)-\frac{1}{2} \left(\boldsymbol \beta_{i}^{\prime}-\boldsymbol \mu^{(t)}\right)^{T}\Sigma^{-1^{(t)}}\left(\boldsymbol \beta_{i}^{\prime}-\boldsymbol \mu^{(t)}\right)\right\}}{\exp \left\{-\frac{1}{2 \sigma^{2(t)}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t)}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t)}\right)-\frac{1}{2}\left(\boldsymbol \beta_{i}^{(t)}-\boldsymbol \mu^{(t)}\right){ }^{T} \Sigma^{-1(t)}\left(\boldsymbol \beta_{i}^{(t)}-\boldsymbol \mu^{(t)}\right)\right\}}
$$
where $\boldsymbol \beta_i^{\prime}$ consisting with $\beta_{j,i}^{\prime}$, $\beta_{k,i}^{(t)}$ for $k>j$ and $\beta_{k,i}^{(t+1)}$ for $k<j$. The log of the ratio is
$$
\begin{aligned}
\log \frac{\pi\left(\boldsymbol \beta_{i}^{\prime},\boldsymbol \Theta_{(-\boldsymbol \beta_i)}^{(t)} \mid \boldsymbol Y\right)}{\pi\left(\boldsymbol \beta_{i}^{(t)}, \boldsymbol \Theta_{(-\boldsymbol \beta_i)}^{(t)} \mid \boldsymbol Y\right)} &=-\frac{1}{2}\left(\frac{\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{\prime}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{\prime}\right)}{\sigma^{2(t)}}+\left(\boldsymbol \beta_{i}^{\prime}-\boldsymbol \mu^{(t)}\right)^{T} \Sigma^{-1^{(t)}}\left(\boldsymbol \beta_{i}^{\prime}-\boldsymbol \mu^{(t)}\right)\right)\\
&+\frac{1}{2}\left(\frac{\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t)}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t)}\right)}{\sigma^{2(t)}}+\left(\boldsymbol \beta_{i}^{(t)}-\boldsymbol \mu^{(t)}\right)^{T} \Sigma^{{-1}^{(t)}}\left(\boldsymbol \beta_{i}^{(t)}-\boldsymbol \mu^{(t)}\right)\right)
\end{aligned}
$$

Then we randomly sample $u$ from $U(0,1)$ and compare $\log(u)$ with the log ratio. If the $\log(u)$ is smaller, we accept $\beta_{j,i}^{\prime} = \beta_{j,i}^{(t+1)}$, otherwise we reject $\beta_{j,i}^{\prime}$ and $\beta_{j,i}^{(t)} = \beta_{j,i}^{(t+1)}$.


Then, Gibb step for $\boldsymbol \mu$: Sample $\boldsymbol \mu^{(t+1)}$ from $\mathcal{N} \left(\bar{\boldsymbol \beta}^{(t+1)} ,\Sigma^{(t)}/m\right)$, where $\bar{\boldsymbol \beta}^{(t+1)}$ is the average $\boldsymbol \beta_i^{(t+1)}$ over all hurricanes.


Next, is the MH step to generate $\sigma^{2^{\prime}}$ from $U\left(\sigma^{2^{(t)}}-a_{\sigma^2},\sigma^{2^{(t)}}+a_{\sigma^2}\right)$. Firstly, check whether $\sigma^{2^{\prime}}$ is positive, if not, we reject $\sigma^{2^{\prime}}$. Then, we randomly sample $u$ from $U(0,1)$ and compare $\log(u)$ with
the log posterior ratio. The log posterior ratio
$$
\begin{aligned}
\log \frac{\pi\left(\sigma^{2^{\prime}},\boldsymbol \Theta^{(t)}_{(-\sigma^{2})} \mid \boldsymbol Y\right)}{\pi\left( \sigma^{2^{(t)}},\boldsymbol \Theta^{(t)}_{(-\sigma^{2})} \mid \boldsymbol Y\right)} 
&= -\left(1+\frac{M}{2}\right)\log(\sigma^{2^{\prime}})-\frac{1}{2\sigma^{2^{\prime}}}\sum_{i=1}^{m}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t+1)}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t+1)}\right)\\
&+\left(1+\frac{M}{2}\right)\log(\sigma^{2^{(t)}})+\frac{1}{2\sigma^{2^{(t)}}}\sum_{i=1}^{m}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t+1)}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}^{(t+1)}\right)
\end{aligned}
$$

where $M$ is total number of observation for all hurricanes. If the $\log(u)$ is smaller, we accept $\sigma^{2^{\prime}} = \sigma^{2^{(t+1)}}$.


Finally, we sample $\Sigma^{-1^{(t+1)}}$ from  $\mathrm{Wishart}\left(3d+3+m,\left(\boldsymbol I+\sum_{i=1}^{m}\left(\boldsymbol \beta_{i}^{(t+1)}-\boldsymbol \mu^{(t+1)}\right)\left(\boldsymbol \beta_{i}^{(t+1)}-\boldsymbol \mu^{(t+1)}\right)^{T}\right)^{-1} \right).$ 

For the MCMC process we will run 10,000 iterations using the code shown in Appendix B. 

## 2.4  Initial Starting Values for MCMC 

To initiate the MCMC process we need to specify the starting values. The choice of starting value is important to help with the convergence time of our algorithm. For each parameter we will chose the start values to be: 
 \begin{itemize}
    \item $\boldsymbol \beta_{i}$: Fit OLS multivariate linear regression (MLR) for $i^{th}$ hurricane and use the coefficients as $\boldsymbol \beta_{i}^{(0)}$
    \item $\boldsymbol \mu$: Average over all $\boldsymbol \beta_{i}^{(0)}$ as $\boldsymbol \mu^{(0)}$ 
    \item $\sigma^2$: $\hat{\sigma}^2_i$ is the mean square residuals of the OLS model for $i^{th}$ hurricane. Take the mean over all $\hat{\sigma}^2_i$ as $\sigma^{2^{(0)}}$
    \item $\Sigma^{-1}$: Generate the covariance matrix of $\boldsymbol \beta_{i}^{(0)}$ and take the inverse of the matrix as $\Sigma^{-1^{(0)}}$
    \end{itemize}

There is a MH step in the MCMC algorithm so the choice of search window is important. The acceptance rate of the MH step is around 0.317 to 0.637, so the search window of our algorithm is appropriate. We tune the search windows multiple times to achieve this result. Table 1 demonstrates the range of search windows $a$ and associated acceptance rates for $\boldsymbol \beta$ and $\sigma^2$.

\begin{table}
  \centering
  \caption{Range of Search Window and Acceptance Rate for paraemters used MH step}
  \begin{tabular}{|c|c|c|} \hline 
   & Search Window& Acceptance Rate (\%) \\ \hline 
      $\boldsymbol \beta_{0,}$ & 1.1 & 45.87 - 51.36\\ \hline 
      $\boldsymbol \beta_{1,}$ & (0.04, 0.1) & 31.67 - 63.68\\ \hline 
      $\boldsymbol \beta_{2,}$ & (0.8, 1.0) & 38.60 - 45.60\\ \hline 
      $\boldsymbol \beta_{3,}$ & (0.5, 0.6) & 33.20 - 61.32\\ \hline 
      $\boldsymbol \beta_{4,}$ & (0.4, 0.5) & 34.95 - 60.45\\ \hline 
      $\sigma^2$ & 2.0 & 44.83\\ \hline 
  \end{tabular} 
\end{table}



## 2.5 MCMC Model Performance 

To assess the quality of the proposed model we will assess the overall adjusted  $R^2$ value from the Bayesian model. In addition the model performance will be assessed for each hurricane by the adjusted $R^2$ value as well as a goodness of fit test. The goodness of fit test we will use residuals of Bayesian estimates, $r_{ij}$, of the $j^{th}$ observation in $i^{th}$ hurricane, to calculate the test statistics. $\chi^2_{stat} = \frac{\sum_{j=1}^{n_i}r_{ij}^2}{\sigma^2}$, where $\sigma^2$ is the estimate $\sigma^2$ from MCMC. Based on the normal assumption in intro, $\chi^2_{stat}\sim\chi^2_{n_i}$, where $n_i$ is the number of observation for $i^{th}$ hurricane. Visual inspection can also be used for each individual hurricane plotting their observed wind speed with the model's predicted wind speed. 

## 2.6 Models to Explore Seasonal Differences and Yearly Trends 

In the dataset information about the hurricanes start month, year and the nature of the hurricane is recorded. It is of interested to explore how these three variables affect the wind speed. Specifically we are interested in exploring the seasonal differences, and if there any evidence supporting the statement that "the hurricane wind speed has been increasing over years". To do visual inspection can first be used to help us understand how these three variables impact the beta estimates. Then we can see if there is a linear trend given the three variables using each $\beta$ estimates gained from the Bayesian model described above. 
For each $\beta$ value we fit a different linear model. We can first fit a model using the three variables (month, year, and nature) as predictors and the 5 different $\beta$ values as the outcome. This will result in 5 different linear models, one for each $\beta$ value. 
$$
Y_{ji} = \alpha_{0j} + \alpha_{1j}\times \textrm{Decade}_i 
+ \alpha_{(k+1)j}I(\textrm{Nature} = k)_i
+ \alpha_{(l+5)j}I(\textrm{Month} = l)_i + \epsilon_{ji}
$$
Where $i$ is the hurricane, $j$ is the Beta model (0 through 4), $k \in (ET,\ NR,\ SS,\ TS)$ making DS the reference group. Let $l$ in (April - December) where January is the reference group. We chose to include nature and month as categorical variables and year (transformed into decade) as a continuous variable. 

We are also going to consider 5 different models just using decade as a predictor.  $Y_i = \alpha_{0i}+\alpha_{1i} \times\textrm{Decade}$ where $Y_i$ is each $\beta_i$ and $i \in (0,\dots,4)$.  


## 2.7 Models to Explore Death and Damages 

\textbf{FILL IN}


# 3. Results 

## 3.1 MCMC Model Convergence 
Using the starting values describe above we can now examine the model performance. We are particular interested in the parameter convergence. Below we can see Figure \ref{fig:converge} that displays each of the parameters of iterations over the 10,000 iterations. Wee see that each parameter reaches a convergence at around 5,500 iterations. Therefore, we will make 5,500 iterations our burn in period. Figure \ref{fig:hist_select} displays the histogram results from the selected parameters excluding the burn in period. We see that each of the distributions is relatively normal with some skewed results for $\Sigma^{-1}$ and heavy tails in $\mu_0$. 

Convergence was reach in each of the parameters of interest. Thus we can take the average estimate excluding the burn in period. When we do this we are left with the $\mu_i$ values, $\sigma^2$ values, $\Sigma^2$ matrix, and a $\beta_i$ estimate for each hurricane. Figure \ref{fig:mu_est} displays the $\mu_i$ values, $\sigma^2$ values, $\Sigma^2$ matrix results.  We see that the $\mu_2$, $\mu_3$ estimate are negative while the others are positive. Figure \ref{fig:beta_est} displays the $\beta_i$ estimates for each hurricane in a histogram. 

The average $\beta_0$ estimate for the hurricanes is 3.80 which represents the intercept of the wind speed model.  The average $\beta_1$ estimate for the hurricanes is 0.915 which represents the lagged wind speed impact model. This positive estimate indicated that a larger previous wind speed will indicate a higher wind speed for the next time point. The $\beta_2$ estimate is -0.136 which represents the impact of the change in latitude. The $\beta_3$ estimate is -0.395 which represents the impact of the change in longitude. Both of the impact of change in latitude and longitude have negative beta estimated values indicating that a larger change in direction corresponds with a decrease in wind speeds.  Thus hurricane speed and wind speed seem to be inversely correlated. The $\beta_4$ estimate is 0.481 which represents the impact of the change in wind speed. This positive value indicated that a larger change in wind speed corresponds with a larger wind speed value. 


\begin{figure}
    \centering
    \includegraphics{pics/covergence_plots.png}
    \caption{Convergence Plot of the Selected Parameters after 10,000 Iterations}
    \label{fig:converge}
\end{figure}


\begin{figure}
    \centering
    \includegraphics{pics/hist_select.png}
    \caption{Histogram Plot of the Selected Parameters not including the 5,500 burn in period results}
    \label{fig:hist_select}
\end{figure}


\begin{figure}
    \centering
    \includegraphics{pics/hist_beta.png}
    \caption{Bayesian Estiamtes for the $\beta$ values for all Hurricane}
    \label{fig:beta_est}
\end{figure}


\begin{figure}
    \centering
    \includegraphics{pics/mu_est.png}
    \caption{Bayesian Estiamtes for $\mu$ and $\sigma^2$}
    \label{fig:mu_est}
\end{figure}



## 3.2 MCMC Model Performance 

```{r message=FALSE, warning=FALSE, include=FALSE}
r_square = (1-sum((dt_res$wind_kt_pred-dt_res$wind_kt)^2)/sum((dt_res$wind_kt-mean(dt_res$wind_kt))^2))
adj_r_square = 1-(1-r_square)*(n_obs-1)/(n_obs-n_beta)

dt_rsquare = dt_res %>% 
  dplyr::select(id, r_square, adj_r_square, p_value, n_obs) %>% 
  distinct() %>% 
  arrange(n_obs)

n_lowp = dt_rsquare %>% filter(p_value < 0.05) %>% 
  nrow()

d_high = dt_rsquare %>% 
  filter(adj_r_square >=0.6 & adj_r_square <= 1) %>% 
  nrow()

d_moderate = dt_rsquare %>% 
  filter(adj_r_square >= 0.2 & adj_r_square< 0.6) %>% 
  nrow()

d_low = dt_rsquare %>% 
  filter(adj_r_square < 0.2) %>% 
  nrow()


```

The overall adjusted $R^2$ of the estimated Bayesian model is `r adj_r_square`. We can also look at the adjusted $R^2$ value for each individual hurricane. Table 2 displays the values grouped by the adjusted $R^2$ showing the number of hurricanes and the percentage. We see that most hurricanes do well with a adjusted $R^2$ above 0.6. Even though most of the estimated models track hurricanes well with great adjusted $R^2$, some estimated models track the hurricanes extremely bad. `r round((d_low+d_moderate)/681,3)*100`% of the estimated models do not track the hurricane well and have adjusted $R^2$ less than 0.6. Few models have negative adjusted $R^2$. We also perform the goodness-of-fit test for individual estimated model. The estimated models of `r n_lowp` hurricanes have p-value less than 0.05, implying that those models do not fit well. 

Figure \ref{fig:adj_r} shows the adjusted $R^2$ for each hurricane by their number of observations. This plot shows that with more observations the better our model with do. The to the right of the vertical line shows that 40 or more observations give us an adjusted $R^2$ values above 0.5.

Figure \ref{fig:pred_id} displays 9 randomly selected hurricanes actual wind speeds and predicted wind speeds. We see that for all of them the lines are very similar with some slight derivations away when the number of observations are small. In DOG 1950 we see very good model prediction and for this particular hurricane we have a lot of observations. However, for JERRY 2007 we do not have as many observations and see deviations at the begging and end of the storm. 


```{r echo=FALSE, message=FALSE, warning=FALSE}

label_1 = c("$R_{adj}^2$", "Count of Hurricanes","Percentage($\\%$)")

tibble(adj_r = c("0.6-1", "0.2-0.6","< 0.2"),
       count = c(d_high, d_moderate, d_low),
       prop = c(round(d_high/681,3)*100, round(d_moderate/681,3)*100, 
                round(d_low/681,3)*100)) %>% 
  knitr::kable(escape = F, col.names = label_1, 
        caption = "$R_{adj}^2$ for each hurricane") 
# %>% 
#   kable_styling(latex_options = "hold_position")

```



\begin{figure}
    \centering
    \includegraphics{pics/adj_r_square.png}
    \caption{Adjusted R squared for each Hurricane plotted by the number of observations.}
    \label{fig:adj_r}
\end{figure}

\begin{figure}
    \centering
    \includegraphics{pics/predict_id_plot.png}
    \caption{The predicted wind speeds verses the actual observed wind speeds for 9 randomly selected hurricanes. }
    \label{fig:pred_id}
\end{figure}



## 3.3 Seasonal Differences and Yearly Trends Results 
Using the model output obtained in the Bayesian MCMC algorithm we can now explore how the hurricane's start month, nature, and year affect the wind speed. Particularly we are interesting in if there are any seasonal difference and if there is evidence to support the statement that "the hurricane wind speed has been increasing over years". 

We can first explore if there are seasonal difference. Visually inspecting the $\beta$ values obtained from the MCMC process with respect to year, month and nature we can look at Figure \ref{fig:vis_inspect_year} and \ref{fig:vis_inspect_nature}. Looking at the beta estimates over the years grouped by month and the betas, Figure \ref{fig:vis_inspect_year}, we can see that the majority of the hurricanes fall within the June through November months. We can also see that looking at December in the later years there seems to be more Hurricanes. This time spend is what is conventionally accepted as hurricane season. Within each month there does not seem to be any obvious change in the beta estimates over the years. Looking at the Nature of the hurricane by the beta's box plot, Figure \ref{fig:vis_inspect_nature}, we see that all the box plots for the nature seem to be relatively similar within each beta estimates. This indicates that there doesn't seem to be drastic differences between the 5 different hurricane natures. However, looking at beta1 section we see that the Extra Tropical (ET) box seems to be much larger than the disturbance (DS), not rated (NR) and sub tropical (SS) nature types. This visually implies that the extra tropical hurricane has increased importance of previous time wind speed impacting the next wind speed prediction. 

Fitting the linear model using decade, month and nature as predictors results in 5 different model, one for each $\beta$ vector. The model output can be seen in Appendix C, due to the length it was not included in the main body. To summarize the results no nature, or month indicators were significant. The $\beta_1$ model had the decade estimates as significant -0.003 (-0.002, -0.004). Thus month and nature don't have a linear association within each beta vector when adjusting for the year. 

Just exploring the decade variable we can address the question of does "the hurricane wind speed has been increasing over years"? To do this we can do a visual inspection as well as a linear association test. Figure \ref{fig:year_model} explore this question by showing a visual inspection on the left and a linear model on the right. The model output just shows the estimates related to the decade using each beta vector as an outcome. The results are the $\beta_1$ and $\beta_3$ decade estimate are the only ones that are significant. 


\begin{figure}
    \centering
    \includegraphics{pics/year_month_dot.png}
    \caption{Exploring the Beta estimates by month and year.  }
    \label{fig:vis_inspect_year}
\end{figure}


\begin{figure}
    \centering
    \includegraphics{pics/nature_box.png}
    \caption{Exploring the Beta estimates by nature of the hurricanes }
    \label{fig:vis_inspect_nature}
\end{figure}


\begin{figure}
    \centering
    \includegraphics{pics/year_model_visual}
    \caption{Exploring how wind speed has changed over the years }
    \label{fig:year_model}
\end{figure}


## 3.5 Death and Damages Results 



# 4. Discussion 


## 4.1. Summary of Findings


## 4.2. Limitations



## 4.3 Future Work



## 4.4. Group Contributions



\newpage

# References 

\begin{enumerate}[label={[\arabic*]}]
  \item  one 
  \item  two 
\end{enumerate}

\newpage 

# Appendices


# Appendix A 

Calculating the conditional Priors for each paramater. 

$$
\begin{aligned}
&\boldsymbol \beta_i:
\pi(\boldsymbol \beta_i \mid \boldsymbol \Theta_{(-\boldsymbol \beta_i)} \boldsymbol Y) \propto \exp \left\{-\frac{1}{2}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)-\frac{1}{2 \sigma^{2}}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\}\\
&\boldsymbol \mu:
\pi\left(\boldsymbol \mu \mid \boldsymbol \Theta_{(-\boldsymbol \mu)}, \boldsymbol Y\right)\propto
\exp \left\{-\frac{1}{2} \sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\right\}
\end{aligned}
$$
$$
\begin{aligned}
(\boldsymbol \beta_{i}-\boldsymbol \mu)^{T} \Sigma^{-1}(\boldsymbol \beta_{i}-\boldsymbol \mu)
&=\operatorname{tr}\left(\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\right)\\
&=\operatorname{tr}\left(\Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)(\boldsymbol \beta_i-\boldsymbol \mu)^{T}\right)\\
\pi\left(\boldsymbol \mu \mid \boldsymbol \Theta_{(-\boldsymbol \mu)}, Y\right)&\propto \exp\left\{-\frac{1}{2}\operatorname{tr}\left(\Sigma^{-1}\sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)(\boldsymbol \beta_i-\boldsymbol \mu)^{T}\right)\right\}\\
&= \exp\left\{-\frac{1}{2}\operatorname{tr}\left(\Sigma^{-1}m\left(\boldsymbol \mu - \bar{\boldsymbol \beta}\right)(\boldsymbol \mu - \bar{\boldsymbol \beta})^{T}\right)\right\}\\
&= \exp \left\{-\frac{1}{2}(\boldsymbol \mu-\bar{\boldsymbol \beta})^{T}\Sigma^{-1} m(\boldsymbol \mu-\bar{\boldsymbol \beta})\right\}\\
&\Rightarrow \boldsymbol \mu \mid \boldsymbol \Theta_{(-\boldsymbol \mu)}, \boldsymbol Y \sim N(\bar{\boldsymbol \beta},\Sigma/m)
\end{aligned} 
$$
$$
\sigma^2: \pi\left(\sigma^{2} \mid \boldsymbol \Theta_{(-\sigma^{2})}, \boldsymbol Y\right) \propto \left(\sigma^{2}\right)^{-\left(1+\frac{\sum_{i=1}^{m} n_{i}}{2} \right)}
\times \exp \left\{-\frac{1}{2 \sigma^{2}} \sum_{i=1}^{m}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)^{T}\left(\boldsymbol Y_{i}-\boldsymbol X_{i} \boldsymbol \beta_{i}\right)\right\}
$$


$$
\begin{aligned}
\Sigma^{-1}:
\pi\left(\Sigma^{-1} \mid \boldsymbol \Theta_{\left(-\Sigma^{-1}\right)},\boldsymbol Y\right)
&\propto
\left|\Sigma^{-1}\right|^{d+1+\frac{m}{2}} \exp \left\{-\frac{1}{2} \operatorname{tr}\left(\Sigma^{-1}\right)\right\}\exp \left\{-\frac{1}{2} \sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T} \Sigma^{-1}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\right\}\\
&=\left|\Sigma^{-1}\right|^{d+1+\frac{m}{2}}\exp \left\{-\frac{1}{2}\operatorname{tr}\left(\Sigma^{-1}\right)-\frac{1}{2}\operatorname{tr}\left(\Sigma^{-1} \sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T}\right)\right\}\\
&=\left|\Sigma^{-1}\right|^{d+1+\frac{m}{2}}
\exp \left\{-\frac{1}{2}\operatorname{tr}\left(\Sigma^{-1}\left(I+\sum_{i=1}^{m}\left(\beta_{i}-\mu\right)\left(\beta_{1}-\mu\right)^{T}\right)\right)\right\}\\
&\Rightarrow \Sigma^{-1} \mid \boldsymbol \Theta_{\left(-\Sigma^{-1}\right)}, \boldsymbol Y \sim \mathrm{Wishart}\left(3d+3+m,\left(\boldsymbol I+\sum_{i=1}^{m}\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)\left(\boldsymbol \beta_{i}-\boldsymbol \mu\right)^{T}\right)^{-1} \right)
\end{aligned}
$$



# Appendix B 
The code associated with the MCMC algorithm. 




# Appendix C 
The Model Estimates from Section 3.3. 

\begin{figure}
    \centering
    \includegraphics{pics/linear_est_all.png}
    \caption{The Linear estimates for each beta vector model to explore the effect of nature, month and decade on the wind speed. This corresponds with section 3.3}
    \label{fig:linear_est_all}
\end{figure}


